\section*{ASCA\_ANL\_contrib}
The ASCA\_ANL working group offers an opportunity
to exchange useful modules created by the ASCA\_ANL users.
The ASCA\_ANL\_contrib is a system
to gather useful modules and useful functions as user's contribution
and to re-distribute them to users.

There are two ways to access the ASCA\_ANL\_contrib.
\begin{enumerate}
\item Anonymous FTP \\
  By courtesy of R.Fujimoto and other people at ISAS,
  we can use a following anonymous FTP servise for the ASCA\_ANL\_contrib.
  \begin{quote}
     {\tt ftp://ftp.astro.isas.ac.jp/asca/software/isas/asca\_anl\_contrib/}
  \end{quote}
  This direcotory can also be seen as
  \begin{quote}
     {\tt ISAS:/xray/s1/ftp/software/isas/asca\_anl\_contrib/}.
  \end{quote}
  Read {\tt README} (in Japanese) first.
\item WWW (World Wide Web) \\
  We also put the same information as a hyper text on WWW.
  You can browse the same information with ``Mosaic''.
  \begin{quote}
     {\tt file://ftp.astro.isas.ac.jp/asca/software/isas/asca\_anl\_contrib/Summary.html}
  \end{quote}
  Read {\tt About ASCA\_ANL\_contrib} (in Japanese) first.
\end{enumerate}

Any contribution on the ASCA\_ANL of you are welcomed.
Your contribution is not necessarily an ASCA\_ANL module.
Usefule function, useful shell script, useful program,
all these are accepted as your contribution
as far as it is related to the ASCA\_ANL.
There are some restrictions to put your contribution to the ASCA\_ANL\_contrib.
It is described in {\tt README} in the anonymous ftp directory
and in {\tt About the ASCA\_ANL\_contrib} on WWW.
Please read it carefully before you put your contribution.

We are looking forward to your contributions.
